---
title: "How Do Environmental Treaties Connect?"
authors: "James Hollway, Esther Peev, and Henrique Sposito"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How Do Environmental Treaties Connect?}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## qEnviron Data

 `{qEnviron}` stores several datasets on various aspects of environmental treaties in different databases. For example, the **Agreements** database is made of multiple datasets of environmental treaty titles with specific information such as the agreement type (e.g. agreements, protocols, amendments, declarations), if they are bilateral or multilateral, and some important dates such as the signature or entry into force. The **Memberships** database contains lists of countries with the treaties to which they are parties to. It also includes information on the signature date of the treaty or the withdrawal date of a party. The **References** database is composed of a dataset that indicates the relationship (e.g. amended by, cites, superseded by) between two treaties. Finally, the **Texts** database stores the text of environmental treaties.

## Connecting Environmental Treaties

There are many different ways in which treaties could be linked. For example, a treaty can be substituted by another, complemented by protocols, changed through amendments, and/or cite another treaty. To facilitate the identification and connection between the databases in the package, the `qCreate::code_agreements()` function has been developed to generate a qID for treaties in a database. Each of the `{qEnviron}` dataset contains a qID column. To understand what are the information provided through by the qID, as well as the benefits of using qID to spot the relationships between the treaties across several datasets in a database, please visit this [article](https://globalgov.github.io/qCreate/articles/agreements.html) on the `{qCreate}` website.

Below are two examples of how to visualise treaty linkages for the references and the memberships databases in `{qEnviron}`. We use `{migraph}` to create the network plots, more information about the package can be found [here](https://snlab-ch.github.io/migraph/).

### Visualising How Agreements Connect by References

The graph below illustrates a sample of earliest 25 treaties from the references database. We select only treaties that cite other treaties. The qIDs are used to facilitate the reading and illustration of the relationships.

```{r references, warning=FALSE, message=FALSE, fig.align='center', fig.asp=1.2, fig.width=7}
library(dplyr)
library(migraph)
library(ggplot2)
references <- qEnviron::references$ECOLEX_REF %>% 
  dplyr::distinct() %>%
  dplyr::mutate(year = stringr::str_extract(qID2, "[:digit:]{4}")) %>%
  dplyr::filter(RefType == "Cites") %>%
  dplyr::arrange(year)
migraph::gglineage(references[1:25,c(1,2,3)]) +
  labs(title = "Treaty Lineage of Selected Agreements",
       caption = "Source: qEnviron") +
  theme(plot.title = element_text(family = "sans",
                                  size = 18,
                                  hjust = 0.5))

# More minimal example attempt with treaty titles; not very pretty...

referencesmin <- qEnviron::references$ECOLEX_REF %>% 
  dplyr::distinct() %>%
  dplyr::mutate(year = stringr::str_extract(qID2, "[:digit:]{4}")) %>%
  dplyr::filter(RefType == "Cites" & qID2 == "RAMSA_1971A") %>%
  dplyr::arrange(year)
agreementinfo <- qEnviron::agreements$ECOLEX %>%
  dplyr::filter(qID == "RAMSA_1971A") %>%
  dplyr::slice(rep(dplyr::row_number(), 3))
agreementinfonames <- qEnviron::agreements$ECOLEX %>%
  filter(qID == "ST04DC_2014P" | qID == "MDA-UKR[SDD]_2012A" |
           qID == "NGWCHA_2008O") %>%
  dplyr::select(Title)
referencesmin$qID2  <- agreementinfo$Title
referencesmin$qID1 <- agreementinfonames$Title

# Plot things
migraph::gglineage(referencesmin) +
  labs(title = "Treaty Lineage of RAMSA",
       caption = "Source: qEnviron") +
  theme(plot.title = element_text(family = "sans",
                                  size = 18,
                                  hjust = 0.5))
```

### Visualising How Agreements Connect by Country Membership

#### In 2D...

The graph below illustrates the connection between parties and treaties in two points in time, 1960 and 1980, using the memberships database. In the graph, countries are represented by the dots, the treaties by the squares, and memberships by the edges. qIDs are also used to refer to treaty titles.

```{r memberships, warning=FALSE, message=FALSE, fig.align='center', fig.asp=1, fig.width=7}
library(ggplot2)
membership_1960 <- qEnviron::memberships$IEADB_MEM %>%
  dplyr::mutate(year = stringr::str_extract(qID_ref, "[:digit:]{4}")) %>%
  dplyr::filter(year == "1960") %>%
  dplyr::select(qID, CountryID, Title) %>%
  migraph::as_tidygraph()
migraph::autographr(membership_1960) +
  labs(title = "Membership in Environmental Treaties\nSigned in 1960",
       caption = "Source: qEnviron") +
  theme(plot.title = element_text(family = "sans",
                                  size = 18,
                                  hjust = 0.5))
membership_1980 <- qEnviron::memberships$IEADB_MEM %>%
  dplyr::mutate(year = stringr::str_extract(qID_ref, "[:digit:]{4}")) %>%
  dplyr::filter(year == "1980") %>%
  dplyr::select(qID, CountryID, Title) %>%
  migraph::as_tidygraph()
migraph::autographr(membership_1980) +
  labs(title = "Membership in Environmental Treaties\nSigned in 1980",
       caption = "Source: qEnviron") +
  theme(plot.title = element_text(family = "sans",
                                  size = 18,
                                  hjust = 0.5))
```

#### ...And in 3D!

Visualizing the network in 3D is a neat way of presenting your data in HTML
documents. This allows you to let the visitors of your site interact with
your network by allowing them to move things around. To get there, we leverage
the [`{threejs}`](https://bwlewis.github.io/rthreejs/) package, an interface
between R and the eponymous JS library.

We start by importing the membership data for agreements signed in 1980 and
set a node attribute `lvl` identifying the set a node belongs to. We move on to
subsetting the graph by only retaining the biggest component and applying the
`layout_as_multilevel()` layout function. We finally make some cosmetic
adjustments before rendering the graph. Note that you can hover over a node
to display its name!

```{r memberships3D, warning=FALSE, message=FALSE, fig.align='center', fig.asp=1, fig.width=7}
# Loading additional packages
library(threejs)
library(graphlayouts)

# Importing things
membership_1980 <- qEnviron::memberships$IEADB_MEM %>%
  dplyr::mutate(year = stringr::str_extract(qID_ref, "[:digit:]{4}")) %>%
  dplyr::filter(year == "1980") %>%
  dplyr::select(qID, CountryID, Title) %>%
  migraph::as_tidygraph()

# Set node attributes "lvl" to separate between countries and agreements
V(membership_1980)$lvl <- if_else(V(membership_1980)$type, 2, 1)

# Subset the graph to only include the biggest component
componentsNeedles <- igraph::components(membership_1980, mode = "strong")
biggest_component_id <- which.max(componentsNeedles$csize)
vert_ids <- V(membership_1980)[componentsNeedles$membership == biggest_component_id]
membership_1980 <- igraph::induced_subgraph(membership_1980, vert_ids)

# Define the 3D layout
membership_1980$layout <- graphlayouts::layout_as_multilevel(membership_1980,
                                                             type = "all",
                                                             alpha = 25,
                                                             beta = 45,
                                                             project2D = FALSE)
V(membership_1980)$color <- c("#014D23", "#36648A")[V(membership_1980)$lvl]
V(membership_1980)$vertex.label <- V(membership_1980)$name

# Render the graph
graphjs(membership_1980,
        main = "Membership in Environmental Treaties Signed in 1980 (Main Component)",
        bg = "white",
        vertex.label = V(membership_1980)$vertex.label,
        vertex.size = ifelse(V(membership_1980)$lvl == 1, 0.8, 2),
        vertex.shape = ifelse(V(membership_1980)$lvl == 1, "square", "sphere"),
        top = 1,
        left = 45,
        font.main = "18px Arial")
```

